---
title: "Intro to R with NYC 311 Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

This is an R Markdown file.

Useful resources for R Markdown and Knitr:
http://rmarkdown.rstudio.com/
https://www.rstudio.com/resources/cheatsheets/

Load packages.

Either require(package.name) or library(package.name) will work. 

```{r}
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate) 
library(stringr)
```

Note the warnings as some functions are masked as one package after the other is loaded. If there is an issue with a masked function, call it by package name: package_name::function().

To get help at any point, type: ?package_name or ?function_name to console to bring up the help.


### Data import:
- read_csv() is part of the readr package, and will import the data as a tibble
- read.csv() is bart of base r and will import the data as a data.frame

```{r}
call.data <- read_csv(file = "311service_2015calls_10per.csv", col_names = TRUE)
call.data <- call.data %>% 
  arrange(desc(`Unique Key`))
call.data
vrbl.names <- str_to_lower(colnames(call.data))
vrbl.names <- gsub(" ", ".", vrbl.names)
colnames(call.data) <- vrbl.names
call.data
rm(vrbl.names)
```

rm(list=ls()) will all variables in the global environment.

```{r}
# which variables have missing values?
sapply(call.data, anyNA)
# how to check how many missing values in one column?
sum(is.na(call.data$closed.date))
sum(is.na(call.data$closed.date)) / nrow(call.data) * 100

# which agencies respond to 311 service calls?
sort(unique(call.data$agency))
length(unique(call.data$agency))
length(unique(call.data$agency.name))
head(unique(call.data$agency.name), n = 20)

call.data %>%
  select(agency:agency.name) %>%
  mutate(combo = paste(agency, agency.name, sep = " / ")) %>%
  group_by(combo) %>%
  summarise(unq = n()) %>%
  #arrange(combo) %>%
  arrange(desc(unq)) %>%
  head(n = 20)
  #ggplot(aes(x = unq)) +
  #geom_histogram(bins = 100)

ggplot(data = call.data, aes(x = agency)) +
  geom_bar(stat = "count")

call.data %>%
  group_by(agency) %>%
  summarise(agency.total = n()) %>%
  arrange(desc(agency.total)) %>%
  ggplot(aes(x = agency, y = agency.total)) +
  geom_bar(stat = "identity")

agency.info <- call.data %>%
  group_by(agency) %>%
  summarise(agency.total = n()) %>%
  arrange(desc(agency.total)) %>%
  mutate(ord.agency = factor(agency, levels = agency))

agency.info %>%
  ggplot(aes(x = ord.agency, y = agency.total)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ylab("Total Number of 311 Calls") +
  xlab("Agency") +
  ggtitle("311 Calls by Agency\nYear: 2015")
```

Where do calls originate?

```{r}
# plot Lat v Long 
call.data %>% 
  ggplot(aes(x = longitude, y = latitude)) + 
  geom_point()
plot(call.data$longitude, call.data$latitude, type = "p")

map.base <- call.data %>%
  ggplot(aes(x = longitude, y = latitude)) +
  coord_quickmap()

map.base +
  geom_point()

map.base +
  geom_point(alpha = 0.2, size = 0.25)

map.base +
  geom_bin2d(bins = 150) +
  scale_fill_viridis(option = "D")

map.base +
  geom_hex(bins = 130) +
  scale_fill_viridis(option = "D")

call.data %>%
  ggplot(aes(x = longitude)) +
  geom_histogram() #+
  #geom_histogram(aes(y = ..density..), alpha = 0.5) +
  #geom_density(color = "red", size = 5)
```

Can we get more useful information relating to location, such as: are certain types of calls more common in different places in NYC?

Six agencies account for 85% of the service calls:

```{r}
agency.info <- agency.info %>%
  arrange(desc(agency.total))
cumsum(agency.info$agency.total) / sum(agency.info$agency.total)
head(agency.info, n = 7)

top.six <- call.data %>%
  filter(agency %in% agency.info$agency[1:6]) %>%
  mutate(agency = factor(agency, levels = agency.info$ord.agency))
length(unique(top.six$agency))
unique(top.six$agency)

top.six %>%
  ggplot(aes(x = longitude, y = latitude, color = agency)) +
  geom_point(alpha = 0.2) +
  coord_quickmap()

top.six %>%
  ggplot(aes(x = longitude, y = latitude)) +
  geom_point(alpha = 0.2, size = 0.25) +
  coord_quickmap() +
  facet_wrap(~agency)
```

What are they typical HPD complaints?

```{r}
hpd.calls <- top.six %>%
  filter(agency == "HPD")
length(unique(call.data$complaint.type))

call.data %>% 
  group_by(complaint.type) %>% 
  summarise(tot = n()) %>% 
  arrange(desc(tot))

hpd.calls %>%
  group_by(complaint.type) %>%
  summarise(tot = n()) %>%
  arrange(desc(tot))

hpd.calls %>%
  group_by(descriptor) %>%
  summarise(tot = n()) %>%
  arrange(desc(tot))
```

How does the type of complaint change over the course of the year?

```{r}
complaints <- call.data %>% 
  group_by(complaint.type) %>% 
  summarise(tot = n()) %>% 
  arrange(desc(tot))

head(call.data$created.date)

pop.comp <- call.data %>%
  filter(complaint.type %in% complaints$complaint.type[1:6]) %>%
  separate(created.date, c("date", "time", "tod"), sep = " ") %>%
  mutate(date = mdy(date))

pop.comp

pop.comp %>%
  group_by(date) %>%
  summarise(tot = n()) %>%
  ggplot(aes(x = date, y = tot)) +
  geom_point()

pop.comp %>%
  select(date, complaint.type) %>%
  group_by(date, complaint.type) %>%
  mutate(number = n()) %>%
  distinct() %>%
  ggplot(aes(x = date, y = number, color = complaint.type, shape = complaint.type)) +
  geom_point(size = 5)
```


http://www.cookbook-r.com/Graphs/
http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html
http://www.sthda.com/english/wiki/ggplot2-essentials

